---
title: "Take-Home Exercise 3: Prototyping geospatial segmentation module for analysis of accessibility in Singapore"
author: "Yung Qi Yang"

format: 
  html:
    code-fold: true
    code-summary: "Reveal Code"
    
execute: 
  eval: true
  echo: true
  warning: false
  
date: "`r Sys.Date()`"

toc: true
toc-location: left
toc-depth: 4

progress: true
---

# Geospatial Segmentation Module

## Preparing the R-Environment

### Importing the required R-packages into the R-Environment

```{r}

pacman::p_load(rgdal, sfdep, tmap, sf, ClustGeo, ggpubr, cluster, factoextra, NbClust, heatmaply, corrplot, psych, tidyverse, GGally)

```

### Importing the geospatial Singapore planning zone data

```{r}

sg_sf <- st_read(dsn = "data/geospatial/Master Plan 2014", layer = "MP14_SUBZONE_WEB_PL") %>% 
  st_transform(crs = 3414)

summary(sg_sf)

sg_sf %>%
  tm_shape() +
  tm_polygons(col = "lightyellow") + 
  tm_compass(type = "rose") + 
  tm_scale_bar() + 
  tm_layout(main.title = "Singapore Planning Subzone 2019")

```

### Importing the aspatial data sets of interest

```{r}

hawker = read_csv("data/aspatial/hawker_coordinates.csv") %>% 
  select(-1)
mrt = read_csv("data/aspatial/MRT_coordinates.csv") %>% 
  select(-1)
school = read_csv("data/aspatial/school_coordinates.csv") %>% 
  select(-1)
supermarket = read_csv("data/aspatial/supermarket_coordinates.csv") %>% 
  select(-1)
mall = read_csv("data/aspatial/mall_coordinates.csv") %>% 
  select(-1)

glimpse(hawker)
glimpse(mrt)
glimpse(school)
glimpse(supermarket)
glimpse(mall)

```

### Transforming the aspatial data sets into simple features, *sf*, objects

```{r}

hawker_sf <- hawker %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 3414)

mrt_sf <- mrt %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 3414)

school_sf <- school %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 3414)

supermarket_sf <- supermarket %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 3414)

mall_sf <- mall %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(crs = 3414)

summary(hawker_sf)
summary(mrt_sf)
summary(school_sf)
summary(supermarket_sf)
summary(mall_sf)

```

## Performing Correlation Analysis of the listed ammenities

### Preparing the ammenity data for correlation computation

```{r}

hawker_count <- st_intersects(sg_sf, hawker_sf) %>% 
  lengths()

mrt_count <- st_intersects(sg_sf, mrt_sf) %>% 
  lengths()

school_count <- st_intersects(sg_sf, school_sf) %>% 
  lengths()

supermarket_count <- st_intersects(sg_sf, supermarket_sf) %>% 
  lengths()

mall_count <- st_intersects(sg_sf, mall_sf) %>% 
  lengths()

ammenities_sf <- sg_sf %>% 
  bind_cols(hawker_count = hawker_count, 
            mrt_count = mrt_count, 
            school_count = school_count,
            supermarket_count = supermarket_count,
            mall_count = mall_count)

ammenities_sf
  
```

### Computing the correlation table

```{r}

cluster_vars.corr = cor(ammenities_sf[, 16:20] %>% 
                          st_drop_geometry())

corrplot.mixed(cluster_vars.corr, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")

```


